---
title: "coherence calibration for model recovery"
author: "ari"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(patchwork)
summary_df <- read.csv('results/results_2025-08-15_summary_naAcc.csv')
```

# simulation code
here are the relevant bits of MATLAB code that I used to generate the data being analyzed in this notebook:
```{matlab}
%%%% in the wrapper script %%%%
nSub = 500; % per cell
nTrial = 1000; % per subject
coherence = 0.51:0.01:0.65;
threshold = 3:3:30;
memoryThinning = 12;
visionThinning = 1;
vizPresentationRate = 1/60;
noisePeriods = 0; 
noNoiseTrialDuration = 3; 

%%%% in the simulation script %%%%
visionEvidence = (binornd(1,coherence, [nFrames*2,nTrial])*2-1) + normrnd(0,1,[nFrames*2,nTrial]);
    noiseFrames = 1:2:size(visionEvidence, 1); % returns the index of every other row
    visionEvidence(noiseFrames, :) = 0; 
    
decisionVariable = cumsum(visionEvidence, 1);
    
% identify trials where DV hit either threshold
mask = abs(decisionVariable) > threshold;

% get index of first threshold passing
[~, rawChoiceRT] = max(mask, [], 1); 
rawChoiceRT = transpose(rawChoiceRT); 

% indicate which trials never made it to threshold
noChoice = transpose(~any(mask, 1));

% set the rawChoiceRTs for those trials as NaN
rawChoiceRT(noChoice) = NaN;

% determine if DV value at threshold crossing corresponds to correct answer (positive value)
rawChoiceAccuracy = NaN(nTrial, 1);
forcedChoiceAccuracy = NaN(nTrial, 1);

for i = 1:nTrial
    if isnan(rawChoiceRT(i))
        forcedChoiceAccuracy(i) = decisionVariable(end, i) > 0;
        rawChoiceAccuracy(i) = 0;
    else
        rawChoiceAccuracy(i) = decisionVariable(rawChoiceRT(i), i) > 0;
        forcedChoiceAccuracy(i) = rawChoiceAccuracy(i);
    end
end
```

This generated a CSV with 75,000,000 trial-level datapoints; 1000 per simulated subject, where each subject was presented with each combination of coherence & threshold. I created a summary dataframe with 75,000 observations, so that each cell is the average over the 1000 trials that each random "subject" contributed. Plotted below are the mean + SE values for each combination of threshold & coherence

# accuracy: raw choice vs. forced choice
```{r, fig.width=12}

p1 <- summary_df %>%
  mutate(threshold = factor(threshold)) %>%
  ggplot(aes(x=coherence, y=rawChoice_accuracy, color=threshold, group=threshold)) +
  theme_bw() + geom_hline(yintercept = 0.5, color='gray30') +
  stat_summary(fun = 'mean', geom='line') +
  stat_summary(fun.data = 'mean_se', geom='pointrange') +
  geom_hline(yintercept = 0.7, linetype = 'dashed') + 
  facet_wrap(~ threshold) +
  labs(title = 'raw choice accuracy') 

## plot accuracy: forced choice
p2 <- summary_df %>%
  mutate(threshold = factor(threshold)) %>%
  ggplot(aes(x=coherence, y=forcedChoice_accuracy, color=threshold, group=threshold)) +
  theme_bw() + geom_hline(yintercept = 0.5, color='gray30') +
  stat_summary(fun = 'mean', geom='line') +
  stat_summary(fun.data = 'mean_se', geom='pointrange') +
  geom_hline(yintercept = 0.7, linetype = 'dashed') + 
  facet_wrap(~ threshold) +
  labs(title = 'forced choice accuracy') 

p1 + p2 + plot_layout(guides = 'collect')
```

very nice curves :) and nice to see the difference in slopes for coherence between raw & forced accuracy. 

because the free-response version is closer to what subjects have to do on the task, I think we should use that way of looking at the data to decide on tuples of coherence & threshold to run in the generating model. 

# RTs for raw choice
here is what RTs look like for the raw choice data:
```{r, echo=F}
summary_df %>%
  mutate(threshold = factor(threshold)) %>%
  ggplot(aes(x=coherence, y=rawChoice_rt, color=threshold, group=threshold)) +
  theme_bw() +
  stat_summary(fun = 'mean', geom='line') +
  stat_summary(fun.data = 'mean_se', geom='pointrange') +
  facet_wrap(~ threshold) +
  labs(title = 'raw choice RT') 
```

# Extracting threshold-level combos of coherence & threshold
```{r}
summary_df %>% group_by(threshold, coherence) %>% summarise(rawChoice = mean(rawChoice_accuracy))

out_df <- summary_df %>%
  group_by(threshold, coherence) %>% 
  summarise(rawChoice = round(mean(rawChoice_accuracy), 3)) %>%
  ungroup() %>%
  group_by(threshold) %>%
  slice_min(abs(rawChoice - 0.7), n = 1, with_ties = F)

out_df
```

```{r, include=F}

write.csv(out_df, '../calibratedCoherences.csv', row.names = F)

```